Constructing Sailing Match Race Schedules: Round-Robin Pairing Lists

### 2015-04-15 paper.tex
I don't think Craig meant to change that
changes by patrick & craig to address craig's comments

Criteron~6 allows us to measure the \emph{balance} of a schedule. Perfect balance would be when a skipper has as many matches in first position in flights as in second position in flights, and so on. For example, if we had 9 skippers and 8 boats, each skipper would have 8 matches and could be in the first, second, third or fourth match of a flight. In perfect balance, each of the 8 skippers would appear twice in each position. Balance is one of our optimisation criteria.

@@

Criterion~10 is ambiguous: this could be interpreted on a schedule basis: i.e.\ to minimise the
overall number of changes in the entire pairing list schedule; or alternatively as well as a fair,
but minimal number of changes across all skippers. In our work, we take the latter viewpoint.

@@

        \begin{scope}[yshift=9.9cm, xshift=-3.5cm]
            \node [anchor=north west] (ModMatchTable) at (0, 0) {
                \begin{minipage}{4.5cm}
                    \centering
                    \setlength{\tabcolsep}{2pt}
                    \setlength{\aboveep}{0pt}
                    \setlength{\belowrulesep}{0pt}
                    \setlength{\extrarowheight}{.75ex}
                    \begin{tabular}{cccccccc}
                        & 0 & 1 & 2 & 3 & 4 & 5 & 6 \\ \cmidrule{2-8}
                        \multicolumn{1}{l@{\hspace{1em}}|}{0} & ~-~ & 0 & 1 & 3 & 5 & 2 & 6 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{1} & 0 & ~-~ & 6 & 2 & 3 & 1 & 4 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{2} & 1 & 6 & ~-~ & 0 & 4 & 5 & 2 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{3} & 3 & 2 & 0 & ~-~ & 6 & 4 & 5 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{4} & 5 & 3 & 4 & 6 & ~-~ & 0 & 1 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{5} & 2 & 1 & 5 & 4 & 0 & ~-~ & 3 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{6} & 6 & 4 & 2 & 5 & 1 & 3 & ~-~ \\
                    \end{tabular}
                \end{minipage}
            };
            \node [above=-0.1 of ModMatchTable.north west, anchor=south west] (ModMatch) { $\modMatch$ };
        \end{scope}

@@

\paragraph{8 skippers and 6 boats:} We first analyse the ISAF schedule, Table~\ref{08-06} on the
left. Criterion~4 (skippers in last match in a flight cannot be first in next flight) is violated on
3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criterion~5 (no
more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3)
and skipper 7 (flights 7 to 9). Criterion~6 (imbalance) is violated for skippers 6 and 8. Criterion~12  (new
boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat
changes. Our schedule, Table~\ref{08-06} on the right, respects all criteria and has 6 boat changes
and a maximum imbalance of 1.

\paragraph{9 skippers:} Two new schedules are presented in Table~\ref{09}, on the left for 6 boats
and on the right for 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both
schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1
boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper
with more than 1 boat change, and each skipper 4 times on starboard side and 4 on port side.

@@

\paragraph{10 skippers and 6 boats:} Table \ref{10-06} shows the ISAF schedule on the left for 10
skippers and 6 boats and on the right, our schedule.  The ISAF schedule violates criterion~5 (no more
than 2 consecutive port or starboard assignments) ten times (twice for skippers 0, 1 and 6 and once
for skippers 3, 7, 8 and 9).  Criterion~12 (new boats do not sail first) is violated 7 times (in
flight 3 for skippers 6 and 2, flight 5 for skipper 8, flight 6 for skipper 7, flight 8 for skipper
1, flight 9 for skipper 4 and flight 12 for skipper 4 again). There are 22 boat changes, with the
minimum changes for a skipper being 1 and the maximum 3. Our schedule, right of Table \ref{10-06},
satisfies all criteria, has 12 boat changes with the minimum changes for a skipper being 0 and the
maximum 2, and a maximum imbalance of 1.

@@

\paragraph{8 skippers and 8 boats (easy-8):} Table \ref{08-08} shows the ISAF schedule on the right
for 8 skippers and 8 boats and on the right, our schedule.  The ISAF schedule violates criterion~4
(last match in a flight cannot be first in next flight) for skipper 5 in flights 1 and 2. Also,
criterion 5 (no skipper should have more than 2 consecutive port or starboard assignments) is
violated for skippers 0, 1, 3, 4 and 6. Furthermore, skipper 1 appears in the last match of four
flights resulting in significant imbalance. Our schedule, on the right, satisfies all criteria, is
optimal (maximum imbalance of 1) and took less than 10 seconds in total (all stages) to produce.

@@

We have only reported a handful of our schedules, however there are missing schedules. Examples of
these missing schedules are 10 skippers and 8 boats, 11 skippers with fewer than 10 boats, 12
skippers and 8 boats, 13 skippers and 8 boats and 14 skippers with fewer than 10 boats. None of these
schedules have been published by the ISAF, although we expect that they can be produced by
selectively violating some of constraints. We have also not yet encoded criterion~7, addressing the
situation of 10 or more boats. To do this we will have to modify the DFA constraint used in the
first two stages.

>>>>

Criteron~6 allows us to measure the \emph{balance} of a schedule. Perfect balance would be when a skipper has as many matches in first position in flights as in second position in flights, and so on. For example, if we had 9 skippers, each skipper would have 8 matches and could be in the first, second, third or fourth match of a flight. In perfect balance, each of the 8 skippers would appear twice in each position. Balance is one of our optimisation criteria, and we do this unconditionally i.e.\ even when there are more skippers than boats.

@@

Criterion~10 is ambiguous: this could be interpreted on a schedule basis: i.e.\ to minimise the
overall number of changes in the entire pairing list schedule; or alternatively as well as a fair,
but minimal number of changes across all skippers. In our work, we minise the overall number of changes.

@@

        \begin{scope}[yshift=9.9cm, xshift=-3.5cm]
            \node [anchor=north west] (ModMatchTable) at (0, 0) {
                \begin{minipage}{4.5cm}
                    \centering
                    \setlength{\tabcolsep}{2pt}
                    \setlength{\aboverulesep}{0pt}
                    \setlength{\belowrulesep}{0pt}
                    \setlength{\extrarowheight}{.75ex}
                    \begin{tabular}{cccccccc}
                        & 0 & 1 & 2 & 3 & 4 & 5 & 6 \\ \cmidrule{2-8}
                        \multicolumn{1}{l@{\hspace{1em}}|}{0} & ~-~ & 0 & 1 & 3 & 5 & 2 & 6 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{1} & 0 & ~-~ & 6 & 2 & 3 & 1 & 4 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{2} & 1 & 6 & ~-~ & 0 & 4 & 5 & 2 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{3} & 3 & 2 & 0 & ~-~ & 6 & 4 & 5 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{4} & 5 & 3 & 4 & 6 & ~-~ & 0 & 1 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{5} & 2 & 1 & 5 & 4 & 0 & ~-~ & 3 \\
                        \multicolumn{1}{l@{\hspace{1em}}|}{6} & 6 & 4 & 2 & 5 & 1 & 3 & ~-~ \\
                    \end{tabular}
                \end{minipage}
            };
            \node [above=-0.1 of ModMatchTable.north west, anchor=south west] (ModMatch) { $\modMatch$ };
        \end{scope}

@@

\paragraph{8 skippers and 6 boats:} We first analyse the ISAF schedule, Figure~\ref{08-06} on the
left. Criterion~4 (skippers in last match in a flight cannot be first in next flight) is violated on
3 occasions (skipper 1 in flight 3, skipper 7 in flight 4, skipper 0 in flight 7). Criterion~5 (no
more than two consecutive port or starboard assignments) is violated for skipper 6 (flights 1 to 3)
and skipper 7 (flights 7 to 9). Criterion~6 (imbalance) is violated for skippers 6 and 8. Criterion~12  (new
boats do not sail first) is violated in flight 9 for skipper 5. Finally, the schedule has 8 boat
changes. Our schedule, Figure~\ref{08-06} on the right, respects all criteria and has 6 boat changes
and a maximum imbalance of 1.

\paragraph{9 skippers:} Two new schedules are presented in Figure~\ref{09}, on the left for 6 boats
and on the right for 8 boats. Neither of these schedules appear in the ISAF Umpires' Manual. Both
schedules respect all criteria. For 6 boats there are 8 boat changes (no skipper making more than 1
boat change) and a maximum imbalance of 2. For 8 boats there are again 8 boat changes, no skipper
with more than 1 boat change, and each skipper 4 times on starboard side and 4 on port side.

@@

\paragraph{10 skippers and 6 boats:} Figure~\ref{10-06} shows the ISAF schedule on the left for 10
skippers and 6 boats and on the right, our schedule.  The ISAF schedule violates criterion~5 (no more
than 2 consecutive port or starboard assignments) ten times (twice for skippers 0, 1 and 6 and once
for skippers 3, 7, 8 and 9).  Criterion~12 (new boats do not sail first) is violated 7 times (in
flight 3 for skippers 6 and 2, flight 5 for skipper 8, flight 6 for skipper 7, flight 8 for skipper
1, flight 9 for skipper 4 and flight 12 for skipper 4 again). There are 22 boat changes, with the
minimum changes for a skipper being 1 and the maximum 3. Our schedule, right of Figure~\ref{10-06},
satisfies all criteria, has 12 boat changes with the minimum changes for a skipper being 0 and the
maximum 2, and a maximum imbalance of 1.

@@

\paragraph{8 skippers and 8 boats (easy-8):} Figure~\ref{08-08} shows the ISAF schedule on the right
for 8 skippers and 8 boats and on the right, our schedule.  The ISAF schedule violates criterion~4
(last match in a flight cannot be first in next flight) for skipper 5 in flights 1 and 2. Also,
criterion 5 (no skipper should have more than 2 consecutive port or starboard assignments) is
violated for skippers 0, 1, 3, 4 and 6. Furthermore, skipper 1 appears in the last match of four
flights resulting in significant imbalance. Our schedule, on the right, satisfies all criteria, is
optimal (maximum imbalance of 1) and took less than 10 seconds in total (all stages) to produce.

@@

We have only reported a handful of our schedules, however there are missing schedules i.e.\ unpublished and not yet produced by us. Examples of
these are 10 skippers and 8 boats, 11 skippers with fewer than 10 boats, 12
skippers and 8 boats, 13 skippers and 8 boats and 14 skippers with fewer than 10 boats. None of these
schedules have been published by the ISAF, although we expect that they can be produced by
selectively violating some of constraints. We have also not yet encoded criterion~7, addressing the
situation of 10 or more boats. To do this we will have to modify the DFA constraint used in the
first two stages. In addition, we have yet to address the format of \emph{two-group} round-robin schedules as also found in \cite{isaf}.
